### 2023-08-30 thesis/sections/4-exp-setup.tex
comments, feedback

In this section I will describe which experiments were conducted to answer the research questions described in the previous chapters.
Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
This is done in three parts.
Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
In chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
Furthermore, these experiments provide the basis for the language games, which are then described in chapter \ref{sec:language-games}.

\subsection{Creation of the dataset}
\label{sec:creation-dataset}
\cmtDK{ambiguity}

This research investigates how agents communicate about the relations of objects seen in images.
For that reason, the original CLEVR dataset offers too little control over how the objects are created and in which relation they stand to each other.
Following, I extended the source code to generate images for the CLEVR dataset.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
To simplify the generation and the succeeding training of the models, I only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
The \emph{material} is always the same for all objects in a generated image.
There were three main extensions to the code:

@@

In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
For example the target object is a \emph{small red cube}.
If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
The number of shared attributes can also be set to a range to make the discrimination task more challenging.

Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
This allows for a very controlled generation of relations between the objects in the scene.
Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
Here, the target object is the large purple cylinder.
The target group contains four objects that share zero to a maximum of two attributes.
It is not controlled, which attributes are shared (they are selected randomly). The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
There are no distractor objects.

For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
The size of the generated images is 480x320 pixels.
10.000 images are created for each of the datasets.
Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.

@@

\subsubsection*{CLEVR color}
The second dataset that is created is called 'CLEVR color'.
The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
For this reason, there exist only two groups in the scene, the target object and distractors.
The distractor group can contain in between 6 and 9 objects.
To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
The distractors have always the same size and shape as the target object, but the color is different.
The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.

As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
All attributes, apart from the color are not discriminating the target object from the distractors.
Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.

\subsubsection*{CLEVR Dale datasets}
The above described dataset is very restrictive on the disambiguating relations between the target object and the rest of the objects.
The number and the type of shared attributes are controlled exactly.
More realistically would be if objects can share a variable number of attributes.
In real situations, there is no restriction at all how objects or things relate to each other.
Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
This emergence of referring attributes and their combination is studied deeper in this work.

For this, I created a dataset that allows almost any relation between a target object and the distractors.
The relations are only restricted by the incremental algorithm for the Generation of Referring Expressions (GRE) described in \citet{Dale1995}.
This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
In other words, the object is described unambiguously using the lowest number of words.
For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
This ensures the uniqueness of the target object.
On the other side, it is not controlled which attributes are shared.
These are assigned randomly.
There is again no control over the relations between distractors, which means that distractors can appear multiple times.

Two datasets following these rules are created.
The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}. The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.

@@

Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
Language games are very complex setups for machine learning models.
The models need to solve multiple tasks at the same time in order to solve the overall problem.
For instance, in a simple setup of a game two agents are involved.
The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
And finally it needs to learn how to identify the target object with this information.
There are many points of possible failure to train the agents.

For this reason, I decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
Mainly, the tasks are I separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
With this division, I can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.

The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
Here, I structured the experiments into three levels.
In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
These combined experiments should lay the basis for how to build up the agents in the language games.

\subsubsection{Coordinate predictor}
\label{sec:coordinate_predictor}
\subsubsection*{Setup}

This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
As described before, the sender should communicate an object in the image and the receiver needs to identify it.
The challenge lies in how the receiver refers to the identified object.
There are multiple possibilities, how it can be done.
One of them could be to describe the target object with human language, using the attributes.
The main goal however is to let the language of the agents emerge as natural as possible.
Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
For this reason, the final task of the receiver will be to 'point' to the target object.
The models are therefore tasked to predict the center coordinates of the target object.
With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.

To achieve this goal, multiple setups of models are tested.
In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
Here the image is first passed through one of the \emph{feature extractors}.
Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
These reduce the dimensions first to 1024 and finally to 2.

@@

With this simple setup, the model is theoretically able to focus on an object in the image.
The problem arises as soon as multiple objects are present in the image.
There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
Therefore, the models need to receive more information.
Here, I try out four different ways to encode and refer to the target object.

\begin{figure}[h]
    \centering
    \subfigure[Model including one-hot encoded attributes and locations]{
        \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
        \label{fig:coordinate_predictor_architecture}
    }
    \subfigure[Model including GRE description]{
        \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
        \label{fig:coordinate_predictor_dale_architecture}
    }
    \caption{Simplified architecture of the coordinate predictors}
\end{figure}

In the first method, I encode the attributes of the target object as \textbf{one-hot encodings}.
There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
These three encodings of the attributes are then concatenated.
The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.

In an extension to this method, I also include the \textbf{center coordinates of all objects} in the image.
This should help the model to identify all possible options to chose from, when predicting the target object.
All the center coordinates are simply extracted and shuffled.
Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
The padded locations consist of two zeros for both coordinates.
For this model, I also made use of a more complex way to encode the image.
Here, I based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
The image is first passed through the feature extractor.
Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
This is then passed again through a final linear layer to predict x- and y-coordinates.

The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citet{Dale1995}.
This opposes the idea described before, to share as few human knowledge as possible with the model.
Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
If the model is able to match parts of the image with human words it would show that the model learned this attribute.
If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
This insight would help for succeeding models that make use of these learnings without human language.

Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
In other words, the object is described unambiguously using the lowest number of words.
The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
This order defines, which attributes can be left out, while still identifying the object uniquely.
This research relies on the following order from most important to least important: shape, color, size
Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
Using all three attributes, this description identifies the object perfectly and uniquely.
Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
This can be taken even one step further by removing also the \emph{color}.
Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.

@@

The model is made up of three parts:
The first part extracts features from the image.
Here, the setup is similar to the previous model.
The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
A \emph{ReLU} function is applied, after both convolutional layers.
As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
In the second part, the caption is encoded, using an LSTM.
Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
The third part is again the predictor of the coordinates of the target object.
Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
Between each linear layer, the \emph{ReLU} function is applied.
This architecture is also inspired by the baselines described in \citet{Johnson2017}.

The forth method, to encode the target object utilizes \textbf{masking} of the image.
For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
The side of the square is always two fifth of the image width, in this case 192 pixels.
This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
The square is filled in white, while the rest of the image is filled in black.
This approach has the advantage of providing as few as possible human bias to the model.
While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
It therefore can only rely on its own extracted visual features when looking at masked images.
The model is presented with both the original image and the masked image.
Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.

The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
This distance needs to be minimized.
The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
It doesn't reflect this change.
For that reason, I also introduced an accuracy score.
For that I defined a fixed size circle with a radius of 20 pixels around the center of each object.
If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
A high accuracy may indicate that the model could identify this specific object better.

>>>>

In this section we will describe which experiments were conducted to answer the research questions described in the previous chapters.
Hereby, the setup of these experiments is detailed, and the results are discussed afterwards.
This is done in three parts.
Chapter \ref{sec:creation-dataset} contains the creation of the datasets that will be used in the experiments.
In Chapter \ref{sec:preexperiments} multiple experiments are conducted to evaluate the created datasets.
Furthermore, these experiments provide the basis for the language games, which are then described in Chapter \ref{sec:language-games}.

\subsection{Creation of the dataset}
\label{sec:creation-dataset}
This research investigates how agents use referring expressions to discriminate objects seen in images based on their relations.
For that reason, the original CLEVR framework offers too little control over how a new dataset is created.
Especially, which objects and in which attributes they share with each other in each image can't be controlled.
Following, we extended the framework for generating new datasets.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}{https://github.com/DominikKuenkele/MLT\_Master-Thesis\_clevr-dataset-gen}}
By this, the objects in the generated images are controlled to have different human-recognizable attributes, namely the \emph{shape}, \emph{size} and \emph{color}.
These attributes also correspond to referring expressions in natural language such as English.
% To simplify the generation and the succeeding training of the models, we only focused on the attributes with a high impact on the appearance of the object, namely the \emph{shape}, \emph{size} and \emph{color}.
The \emph{material} is always the same for all objects in a generated image.
There were three main extensions to the framework:
% SD: Too informal
% SD: We create a new dataset based on the CLEVR framework where we control the appearance of scenes in a referring expression task.
% SD: The scenes are controlled by human-recognisable attributes of object such as object shape, colour and type. These attributes also correspond to referring expressions in natural language such as English.
% SD: We create two contexts of scenes, one with two objects and one with five. 
% SD: We also control for the number of attributes shared between the target and the distractor.
% DK: creation of dataset later (here only the general rules), (added; done)

@@

In a second step, when generating the images it is possible to define the relations between \emph{target object}/\emph{target group} and \emph{target object}/\emph{distractors}.
The relation is defined as \textbf{how many} attributes of the target object are identical with the attributes of a single object in the target group and distractors respectively.
For example the target object is a \emph{small red cube}.
If two attributes are shared between target object and target group, objects in the target group could include \emph{small blue cube}, \emph{big red cube} or \emph{small red sphere}, but couldn't include another \emph{small red cube} or a \emph{small blue cylinder}.
The number of shared attributes can also be set to a range to control how challenging the referring task is.
% SD: To control how challenging the referring task is
% DK: (done)

Lastly, it is also possible to define exactly \textbf{which} attributes should be shared between the target object and the groups.
For example, it can be defined to have the same size for objects in the target group, but have different, randomly selected shapes and colors.
This allows for a very controlled generation of relations between the objects in the scene.
% SD: Introduce other images earlier?
% DK: I think 'Dale-5' is the most interesting image to make the point here. The others are introduced later, in the order of the pictures. So I don't think I also can change the order the pictures. (done)
Figure \ref{fig:clevr-dale-5} shows one generated image with this extended source code.
Here, the target object is the large purple cylinder.
The target group contains four objects that share zero to a maximum of two attributes.
It is not controlled, which attributes are shared (they are selected randomly). The large purple cylinder shares the same color and size with the large purple sphere, the same size with both cubes and no attribute with the small turquoise sphere.
There are no distractor objects.
% SD: It depends what you mean by distractor objects. I’d say they are all distractor objects by the virtue they share attributes. The system has to learn understanding of an intersection of these attributes.
% DK: I refer to the definitions in the paragraphs above, namley target object, target groub and distrctors. (done)

For all generated datasets in the following sections, the general constraints and settings are as close as possible to the original CLEVR dataset.
The size of the generated images is 480x320 pixels.
10.000 images are created for each of the datasets.
Each image contains a maximum of 10 objects, that are not intersecting, have the same minimum distance between objects and are at least partially visible from the camera.
% SD: Note that since objects are generated as a part of 3-d scenes, they may appear differently on images: size, occlusion, shading, rotation. We also expect the model to learn from this noise and which therefore makes the task much harder and a task that approximates natural environment compared to the task where we would use abstract geometric shapes projected on a 2-d plane.
% DK: added in the general description of CLEVR since it is not my extention (done)
% TODO: we create the follwoing dataset ....

@@

\subsubsection*{CLEVR color}
The second dataset that is created is called 'CLEVR color'.
The purpose of this dataset is to create scenes, where the target object is completely unique and as easily identifiable as possible.
For this reason, there exist only two groups in the scene, the target object and distractors.
The distractor group can contain in between 6 and 9 objects.
To make the discrimination as simple as possible the target object and the objects in the target group share exactly two attributes.
% SD: Wouldn’t it make sense to introduce the discussion of how scenes can be generated with attributes somewhere here? Again, I felt we were projecting forward earlier without understanding the task.
% DK: before, I only wanted to outlie the rules for the creation, i.e. what is possible with it. Since the code to create the datasets is also a contribution, I wanted to let it stand on its own. Here I only describe how these rules are applied for each new dataset. (QUESTION)
Furthermore, to simplify the relation between target object and distractors over the whole dataset, it is also controlled which attributes are shared.
The distractors have always the same size and shape as the target object, but the color is different.
The reason for choosing the color as the only discriminating attribute is that it is assumed that the color is easier to learn for neural models as opposed to for instance abstract shapes.
% SD: But why not test the limits and the harder examples?
% DK: In the experiments, the models often hda problems to learn anything (with the complex dataset), which is why I reduced the complexity with the 'color' dataset. (done)

As seen in Figure \ref{fig:clevr-color}, the \emph{small brown cylinder} is unique.
By this, it is possible to refer to the target object using the attributes with four different combinations: the \emph{brown} object, the \emph{brown cylinder}, the \emph{small brown} object and the \emph{small brown cylinder}.
All attributes, apart from the color are not discriminating the target object from the distractors.
Notice as well that this restriction doesn't apply to the distractors, where multiple objects with the same color are allowed.
In other words, the choice of attributes is random for the distractors, and they may overlap.
% SD: In order words, the choice of attributes is random for distractor and there may be an overlap between the same distractor objects. This is because we do not control overlap attributes over distractor, I.e. random.
% DK: (added; done)

\subsubsection*{CLEVR Dale datasets}
% SD: In the previous dataset the target and distractor are discriminated by ONE attribute exactly.
% DK: TODO
The above described dataset is very restrictive in the relation between the objects, where only \emph{one} attribute is used to disambiguate them.
The number and the type of shared attributes are controlled exactly.
In the real world, objects have overlapping attributes and hence objects can only be identified by an intersection of multiple attributes.
In real situations, there is no restriction at all how objects or things relate to each other.
Natural language emerged that can refer to distinct attributes of these objects to discriminate them from each other.
This emergence of referring attributes and their combination is studied deeper in this work.
% SD: In real world, objects have over-lapping attributes and hence a single object can only be identified by an intersection of attributes.
% DK: (rephrased; done)

% SD: However, we do not do this randomly but ge5 inspiration from the Dale and Reiter generation algorithm who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy. This way we approximate the information in the scenes to human cognition and we hope that the system will be able to exploit that (we would really need to study and compare it with another dataset where attributes are add3d randomly to confirm that there is an effect of h7man. Ignition).
% DK: see below (done)
For this, we created a dataset that allows almost any relation between a target object and the distractors.
However, the creation is inspired by incremental algorithm for the Generation of Referring Expressions (GRE) described in \citep{Dale1995} who observe that attributes in descriptions occur in certain order and are added incrementally in a certain hierarchy.
This algorithm ensures that every scene contains a unique object in respect to its and the distractors' attributes.
Using the algorithm, one can refer to an object using its attributes to discriminate it from all other objects as efficiently as possible.
In other words, the object is described unambiguously using the lowest number of words.
For the dataset that means that zero, one or two attributes can be shared between the target object and distractor objects.
This ensures the uniqueness of the target object.
On the other side, it is not controlled which attributes are shared.
% SD: The sharing of attributes should be according to the Dale and Reiter hierarchy, shouldn’t it? Two attributes, big blue (cube|sphere)
% DK: During the creation of the dataset, only the number of shared attributes is shared that we get different numbers of overlaps. Only the captions (that are created later and used as input or output for some models/agents) rely directly on the algorithm. At this point, I only wanted to bring up the reason for this, which is the usage of the algorithm in human referring expressions and how it compares to machine referring expressions. (done)
These are assigned randomly.
There is again no control over the relations between distractors, which means that distractors can appear multiple times.

Two datasets following these rules are created.
The Dale-2 dataset contains one target object and one distractor (see Figure \ref{fig:clevr-dale-2}), while the Dale-5 dataset contains one target object and exactly four distractors.
Consider Figure \ref{fig:clevr-dale-5}, with the target object being the \emph{large purple cylinder}. The large purple sphere shares the size and color, the two cubes only share the size, and the small turquoise sphere doesn't share any attribute.
% SD: It’s actually the reverse in which descriptions are generated, we start left to right: large purple (are shared), the last attribute must thus be unique. Generation proceeds in the opposite order, one would just say sphere. We should clarify this. Now it is very hard to understand.
% DK: That's true for the generation of the descriptions. This is explained in chapter 4.2.1 where the captions are needed for the training as input. As said before, the algorithm was not directly used and is just the reason. (done)

@@

Secondly, the experiments in this chapter provide the basis for the setup of the agents in the language games.
Language games are very complex setups for machine learning models.
The models need to solve multiple tasks at the same time in order to solve the overall problem.
For instance, in a simple setup of a game two agents are involved.
The first agent, the sender, is shown a scene with objects and needs to communicate one target object to the other agent, the receiver.
The receiver is shown the same scene and needs to identify the target object with respect to the message of the sender.
In this case, the sender first needs to learn to encode the scene, all objects and their attributes, as well as the information about the target object into its own game specific space.
In a next step it needs to learn how to translate this encoding into a message that is sent to the receiver.
The receiver then needs to learn to decode this message, after which it needs to learn how to combine the decoded message with its own encoding of the scene and objects.
And finally it needs to learn how to identify the target object with this information.
There are many points of possible failure to train the agents.
% SD: This should go at the beginning where language games with neural agents are introduced.
% DK: TODO

For this reason, we decided to divide the main problem and let the models learn simpler subtasks and increase the complexity step by step.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
This will give a very detailed overview where the models struggle to learn and in which ways they can be improved.
Mainly, the tasks are separated into language games with two agents and classical machine learning tasks without any communication, namely only one 'agent' that solves the task alone.
With this division, we can analyze the learning of the encodings of the scenes separately from the learning of producing and decoding messages.

% TODO maybe move to front...
The final objective of this thesis is to find out, how agents can communicate about relations of objects based on their attributes, namely how to generate and understand referring expressions.
Because of that, the first experiments focus on extracting information from images and combining them with structured knowledge about the objects.
Here, we structured the experiments into three levels.
In the first level, the models are trained to learn the position of objects in the image and attend to specific regions of the image, by understanding a referring expression of the target object.
% SD: This is a natural language understanding or reference resolution task.
% DK: TODO maybe put into methods...
In the second level, the models are trained to differentiate objects in the scene from each other, again by understanding referring expressions.
% SD: Object identification task
% DK: TODO
In the last level, the models are trained to generate referring expressions, more specifically the models learn to caption and describe objects in the image.
% SD: Referring expression generation task
% DK: TODO
These combined experiments should lay the basis for how to build up the agents in the language games.
% SD: We are validating the dataset on 3 tasks
% DK: TODO

\subsubsection{Coordinate predictor}
% SD: Reference resolution task
% DK: TODO
\label{sec:coordinate_predictor}
\subsubsection*{Setup}

This level should help to analyze, how the final task of the language game should look like, in especially what the receiver is tasked to predict.
As described before, the sender should communicate an object in the image and the receiver needs to identify it.
The challenge lies in how the receiver refers to the identified object.
There are multiple possibilities, how it can be done.
One of them could be to describe the target object with human language, using the attributes.
The main goal however is to let the language of the agents emerge as natural as possible.
Including human knowledge into the task would bias also the emerged language towards attributes and words, used in natural language.
For this reason, the final task of the receiver will be to 'point' to the target object.
The models are therefore tasked to predict the center coordinates of the target object.
% SD: Reference resolution with identification of location. The other task, object identification is also resolving reference but it is easier as you are only picking objects. The first task also involves spatial knowledge.
% DK: TODO
With this approach, the models receive few human knowledge, but are still able to rely on all information present in the image to discriminate the objects.

To achieve this goal, multiple setups of models are tested.
In the simplest setup, the model receives only the image \cmtDK{preprocess image} as an input and produces two numbers as an output, the predicted x- and y-coordinate of the target object.
% SD: What are the features?
% SD: Note that these are visual features and not spatial features. It is true that visual features also encode some spatial information, how visual features relate to each other, but such information is very different from the spatial information required to predict coordinates in a coordinate frames, and hence we expect the task will be very hard.
% DK: TODO
Here the image is first passed through one of the \emph{feature extractors}.
Next, the extracted feature vector is flattened and passed through two linear layers with a \emph{ReLU} non-linearity in between.
These reduce the dimensions first to 1024 and finally to 2.

@@

With this simple setup, the model is theoretically able to focus on an object in the image.
% SD: Very precisely. This is higher resolution that attention in the visual models that operates on o 7x7 grid, normally.
% DK: TODO
The problem arises as soon as multiple objects are present in the image.
There is no information available for the model to understand which one of these objects is the actual target object, except for the final calculation of the loss.
Since there is not necessarily a pattern for which object in the image is the correct target object over the whole dataset, the models will likely fail to generalize.
Therefore, the models need to receive more information.
Here, we try out four different ways to encode and refer to the target object.

\begin{figure}[h]
  \centering
  \subfigure[Model including one-hot encoded attributes and locations]{
    \includegraphics[width=0.35\linewidth]{figures/arch_coordinate_predictor.png}
    \label{fig:coordinate_predictor_architecture}
  }
  \subfigure[Model including GRE description]{
    \includegraphics[width=0.315\linewidth]{figures/arch_coordinate_predictor_dale.png}
    \label{fig:coordinate_predictor_dale_architecture}
  }
  \caption{Simplified architecture of the coordinate predictors}
  % SD: How are these encoded?
  % DK: TODO
\end{figure}

In the first method, we encode the attributes of the target object as \textbf{one-hot encodings}.
There is a three-dimensional vector encoding the \emph{shape}, an eight-dimensional vector encoding the color and a two-dimensional vector encoding the two different sizes.
The values of each dimension of these vectors can either be zero or one, depending on the attributes of the target object.
These three encodings of the attributes are then concatenated.
The image is again passed through a feature extractor, before the flattened vector is reduced to 2048 dimensions with a linear layer.
The result is concatenated with the encoded attributes and passed to a final linear layer to predict the coordinates.

In an extension to this method, we also include the \textbf{center coordinates of all objects} in the image.
% SD: Encoded locations?
% DK: TODO
This should help the model to identify all possible options to chose from, when predicting the target object.
All the center coordinates are simply extracted and shuffled.
Since there are varying numbers of objects in the image, this vector of variable length is padded to the maximum number of objects in the dataset.
The padded locations consist of two zeros for both coordinates.
For this model, we also made use of a more complex way to encode the image.
Here, we based the approach on code that implements baselines for the CLEVR dataset \citep{Johnson2017}.
The image is first passed through the feature extractor.
Afterwards, a 2-dimensional convolutional layer, reduces the channels from 2048 to 512 channels with a kernel size of 1.
After applying the \emph{ReLU} function, the resulting matrix is max pooled over two dimensions with both a kernel size and a stride of 2.
The resulting matrix is flattened and concatenated with both the attribute encodings and the shuffled coordinates of all objects.
% SD: Why shuffled? If we order them the way they appear in the image, the model will have more information that they are sequentially related.
% DK: TODO
This is then passed again through a final linear layer to predict x- and y-coordinates.
% SD: Convolutions are only applied on visual features, not object attribute features and locations.
% DK: TODO

The third method encodes the attributes of the target object with human language using the \textbf{incremental algorithm for the Generation of Referring Expressions (GRE)} described in \citep{Dale1995}.
This opposes the idea described before, to share as few human knowledge as possible with the model.
Still, this approach can help to understand and analyze if the model was able to extract information about the objects and more specifically their attributes from the image.
% SD: The question you are asking is whether it is possible to predict location from the visual appearance of the object. This is highly complex task as it requires quite two step reasoning, identification of features with that attribute, e.g. blue and then locating that feature in the image.
% DK: TODO
If the model is able to match parts of the image with human words it would show that the model learned this attribute.
If the model in a next step can learn this for the whole dataset, this would mean that it could generalize over these attributes and assign them to certain regions in an image.
This insight would help for succeeding models that make use of these learnings without human language.

Using the algorithm, one can describe an object using its attributes to discriminate it from other objects as efficiently as possible.
In other words, the object is described unambiguously using the lowest number of words.
The algorithm assumes that there is an order of importance for attributes, such as shape, color and size.
This order defines, which attributes can be left out, while still identifying the object uniquely.
This research relies on the following order from most important to least important: shape, color, size
Given for example the scene from \ref{fig:clevr-dale-5} with the target object being the \emph{big purple cylinder}.
Using all three attributes, this description identifies the object perfectly and uniquely.
Following the algorithm, we could make the description shorter by removing the least important attribute \emph{size} without loosing unambiguity, describing it as the \emph{purple cylinder}.
This can be taken even one step further by removing also the \emph{color}.
Describing it as the \emph{cylinder} still doesn't describe any distractor, since the target object is the only cylinder in the scene.
% SD: See my point earlier about the reversed algorithm between the two tasks 
% DK: TODO

@@

The model is made up of three parts:
The first part extracts features from the image.
Here, the setup is similar to the previous model.
The image is passed through a feature extractor, before it is passed through two 2d convolutional layers, both reducing the channels to 128 with a kernel size of 1.
A \emph{ReLU} function is applied, after both convolutional layers.
As before, the resulting vector is pooled, using max pooling with both a kernel size and stride of 2.
In the second part, the caption is encoded, using an LSTM.
Here, the learned embeddings of each token are parsed by the LSTM and its final hidden state is then used as a summary of the complete caption.
The third part is again the predictor of the coordinates of the target object.
Both the processed image and the final hidden state of the LSTM are flattened and concatenated.
The resulting vector is passed through three linear layers reducing to 1024, 1024 and 2 dimensions respectively.
Between each linear layer, the \emph{ReLU} function is applied.
This architecture is also inspired by the baselines described in \citep{Johnson2017}.
% SD: \citep and \citet Use Name (year) when you are referring to particular person and (Name, year) when you are referring to a paper. Hence, in this case it would be the latter.
% DK: done

The forth method, to encode the target object utilizes \textbf{masking} of the image.
% SD: Oh, before we talked about 3 methods
% DK: done
For this, the image is separated into a fixed size squared area containing the target object and the rest of the image.
The side of the square is always two fifth of the image width, in this case 192 pixels.
This size always includes the whole object if it is large or small, while not being cut off if the target object is too close to the border of the image.
The square is filled in white, while the rest of the image is filled in black.
This approach has the advantage of providing as few as possible human bias to the model.
While even the one-hot encodings contain human knowledge by explicitly encoding human chosen attributes, masking the image will only point the model towards the target object without giving more information.
It therefore can only rely on its own extracted visual features when looking at masked images.
% SD: The information in the masked image is still not enough to identify the precise location, only the region. Hence, it corresponds on attention in the attention models. But those are generating labels and not predicting coordinated. The task is still challenging.
% DK: TODO
The model is presented with both the original image and the masked image.
Both are passed through a feature extractor and afterwards passed through a linear layer, reducing the dimensions to 2048.
The resulting vectors are concatenated and passed through another linear layer that predicts the coordinates.

The test dataset is again evaluated on the euclidean distance of the predicted coordinates to the ground truth coordinates.
This distance needs to be minimized.
The mean of all calculated distances is calculated across the whole epoch, which results in a mean distance score per epoch.
Since this score only takes the average of all predictions into account it doesn't show how every prediction fared individually.
If for instance the prediction of one object is getting more precise with growing number of epochs, but the precision of another object gets worse, the mean distance will stay the same.
% SD: Yes that’s correct but through several epochs we hope we will refine the distance and standard deviation of the error. It should level out. It is not a problem. What you do with a circle and accuracy is that you make the task easier as your pointer is now not pointing to a point but to a larger area.
% DK: TODO
It doesn't reflect this change.
For that reason, we also introduced an accuracy score.
For that we defined a fixed size circle with a radius of 20 pixels around the center of each object.
If the model's prediction lies in this circle, it will be counted as a correct prediction, if it lies outside, it is a false prediction.
These scores are averaged for the epoch and result in an accuracy score, where 100\% means that all predictions were very close to the center coordinates and 0\% means that no predictions were close to the center coordinates.
% SD: But here the score will face the same problem with steward deviation. Hence the only difference here is that pointing is less precise.
% DK: TODO
This of course doesn't give a perfect representation since the size of the objects varies, but it will still show, how precise each individual prediction is.
A high accuracy may indicate that the model could identify this specific object better.
